Formation and evolution of intermediate-mass ratio inspirals.

### 2020-07-13 main.tex
Final 13 july 2020

{\bf
We simulate the evolution of the innermost region of the cluster, modelling the IMBH and two stellar mass BHs as live particles and
the remaining cluster as an external static potential. In order to explore the parameter space, we create 7 different models, each one consisting of 4000 simulations gathered in four different sub-classes, depending on the IMBH mass, for a total of \manuel{28,000} simulations. Our model is sketched in Figure \ref{f1}.

@@

These results have two main implications. \manuel{
Since the IMRI secondary mass reflects the underlying BH mass spectrum, this quantity 
encodes information on the population of BHs lurking in the host cluster. If dynamics had enough time to affect BHs -- e.g. causing the ejection of the most massive via strong scatterings -- or the reservoir of BHs have diminished during the IMBH build-up, the IMRI companion mass distribution will tell us how strongly dynamical processes affected BHs. If, instead, dynamics did not play a major role in shaping the IMBH growth, we would expect that the mass spectrum of IMRI companion would resemble the typical mass spectrum of BHs forming out of a standard stellar initial mass function \citep[e.g.][]{kroupa01}.}

@@

Assessing the retention probability for IMBHs is crucial to place constraints on their possible presence in star clusters. To explore this aspect, we use our models to create a catalogue of IMRI mergers proceeding on a step-by-step basis as follows: 
\begin{enumerate}
\item we divide the IMBH mass range -- $M_\ibh = 10^2-5\times 10^5\Ms$ -- in 15 values evenly distributed in logarithmic values;
\item for each IMBH mass, we create a sample of 100 stellar BHs with masses evenly drawn in the $M_\bh = 3.5-80 \Ms$ range. \manuel{This conservative choice allows us to explore from a general perspective the possible ejection of the IMBH via GW recoil and cope with our ignorance about the actual mass spectrum of stellar BHs involved in an IMRI.};
\item we associate to the IMBH and BH a spin randomly drawn between 0-1;
\item we use \cite{jimenez17} numerical relativity fitting formulae to calculate the IMRI merger remnant mass $M_r$, spin $S_r$, and recoil velocity $v_r$, following the procedure depicted in \cite{arca20}.
\end{enumerate}
For each $M_\ibh -m_\bh$ pair, we repeat steps 3 and 4 100 times and, for each of them, we compare the recoil velocity with the escape velocity from the cluster centre, which we set to $\sigma = 5$ or $50$ km/s.

@@

\manuel{As in the previous section, for each merger event we calculate the merger remnant mass, spin, and recoil kick using the numerical relativity fitting formulae described in \cite{jimenez17} following \cite{arca20}.} Figure \ref{fig:f11a} shows the value of the remnant IMBH spin as a function of the IMRI mass ratio.

@@

To further investigate this point, we explore the effects of a series of subsequent mergers on the IMBH final spin. We assume $M_\ibh = 10^3\Ms$, i.e. a mass such to ensure a retention probability of at least $50\%$ in an IMRI merger event involving a BH smaller than $40\Ms$ (see Figures  \ref{fig:survival} and \ref{fig:f11a}). We set the IMBH spin to either $S_\ibh = 0.1-0.5-0.9$. For each spin value, we model ten consecutive mergers with stellar BHs with masses $3-60\Ms$ and spin drawn from a Gaussian distribution with average $S_\bh = 0.5$ and dispersion 0.1. The resulting merger sequence represents an {\it IMBH merger tree}. For each value of the spin we thus create 100 merger trees and calculate the IMBH spin at each successive merger as shown in Figure \ref{fig:f11b}.
We note that all models tend to converge toward the \manuel{same region  of the spin-mass plane, likely close to the spin value adopted for stellar BHs}, thus suggesting that it could be possible to link the IMBH spin with the overall spin distribution of stellar BHs inhabiting the star cluster. In fact, complementing a measure of the IMBH spin with the spin distribution inferred from stellar BH mergers observed as GW sources can help in unravelling the IMBH evolution, whether it was driven by mergers with compact objects or by stellar feeding.
}

>>>>

{\bf
We simulate the evolution of the innermost region of the cluster, modelling the IMBH and two stellar mass BHs as live particles and
the remaining cluster as an external static potential. In order to explore the parameter space, we create 7 different models, each one consisting of 4000 simulations gathered in four different sub-classes, depending on the IMBH mass, for a total of {\bf 28,000} simulations. Our model is sketched in Figure \ref{f1}.

@@

These results have two main implications. {\bf
Since the IMRI secondary mass reflects the underlying BH mass spectrum, this quantity 
encodes information on the population of BHs lurking in the host cluster. If dynamics had enough time to affect BHs -- e.g. causing the ejection of the most massive via strong scatterings -- or the reservoir of BHs have diminished during the IMBH build-up, the IMRI companion mass distribution will tell us how strongly dynamical processes affected BHs. If, instead, dynamics did not play a major role in shaping the IMBH growth, we would expect that the mass spectrum of IMRI companion would resemble the typical mass spectrum of BHs forming out of a standard stellar initial mass function \citep[e.g.][]{kroupa01}.}

@@

Assessing the retention probability for IMBHs is crucial to place constraints on their possible presence in star clusters. To explore this aspect, we use our models to create a catalogue of IMRI mergers proceeding on a step-by-step basis as follows: 
\begin{enumerate}
\item we divide the IMBH mass range -- $M_\ibh = 10^2-5\times 10^5\Ms$ -- in 15 values evenly distributed in logarithmic values;
\item for each IMBH mass, we create a sample of 100 stellar BHs with masses evenly drawn in the $M_\bh = 3.5-80 \Ms$ range. {\bf This conservative choice allows us to explore from a general perspective the possible ejection of the IMBH via GW recoil and cope with our ignorance about the actual mass spectrum of stellar BHs involved in an IMRI.};
\item we associate to the IMBH and BH a spin randomly drawn between 0-1;
\item we use \cite{jimenez17} numerical relativity fitting formulae to calculate the IMRI merger remnant mass $M_r$, spin $S_r$, and recoil velocity $v_r$, following the procedure depicted in \cite{arca20}.
\end{enumerate}
For each $M_\ibh -m_\bh$ pair, we repeat steps 3 and 4 100 times and, for each of them, we compare the recoil velocity with the escape velocity from the cluster centre, which we set to $\sigma = 5$ or $50$ km/s.

@@

{\bf As in the previous section, for each merger event we calculate the merger remnant mass, spin, and recoil kick using the numerical relativity fitting formulae described in \cite{jimenez17} following \cite{arca20}.} Figure \ref{fig:f11a} shows the value of the remnant IMBH spin as a function of the IMRI mass ratio.

@@

To further investigate this point, we explore the effects of a series of subsequent mergers on the IMBH final spin. We assume $M_\ibh = 10^3\Ms$, i.e. a mass such to ensure a retention probability of at least $50\%$ in an IMRI merger event involving a BH smaller than $40\Ms$ (see Figures  \ref{fig:survival} and \ref{fig:f11a}). We set the IMBH spin to either $S_\ibh = 0.1-0.5-0.9$. For each spin value, we model ten consecutive mergers with stellar BHs with masses $3-60\Ms$ and spin drawn from a Gaussian distribution with average $S_\bh = 0.5$ and dispersion 0.1. The resulting merger sequence represents an {\it IMBH merger tree}. For each value of the spin we thus create 100 merger trees and calculate the IMBH spin at each successive merger as shown in Figure \ref{fig:f11b}.
We note that all models tend to converge toward the {\bf same region of the spin-mass plane, likely close to the spin value adopted for stellar BHs}, thus suggesting that it could be possible to link the IMBH spin with the overall spin distribution of stellar BHs inhabiting the star cluster. In fact, complementing a measure of the IMBH spin with the spin distribution inferred from stellar BH mergers observed as GW sources can help in unravelling the IMBH evolution, whether it was driven by mergers with compact objects or by stellar feeding.
}
